---
title: "Intro to R Markdown"
output: html_document
---


## R Markdown Basics

This is an R Markdown file (.Rmd file extension). It allows you to both do data analysis with R *and* write up the process, results, etc. The file can be compiled into a nice-looking document, great for sharing results (and submitting assignments). For now, we'll knit to HTML documents for simplicity. Eventually, we'll knit to PDF format as well.



The main components of an R Markdown file are the header (at the very top), markdown text (regular writing like this paragraph), and code chunks (where we write R code to conduct analyses). Here is a code chunk example:

```{r}
# lines that start with a '#' are comments, used to describe code

# create an object called mean_1_5, which is the mean of the integers 1-5:
mean_1_5 <- mean(1:5)

# output the newly-created mean_1_5 object:
mean_1_5

# To evaluate/run the code in this chunk, put your cursor inside it and use one of the following keyboard shortcuts: Ctrl+Shift+Enter on a PC or Command+Shift+Enter on a Mac.
```

Notice how the code chunk is separated from the main markdown text: it starts on a new line, beginning with three acute accent symbols (the key next to the 1 on your keyboard). This is followed directly by curly braces with an 'r inside. Actual R code begins on the next line, and the code chunk is completed by typing another three acute accents. 

One way to insert a new R code chunk is by clicking the insert button (it has a green square with a white 'c' inside) then 'R'. You should get used to the keyboard shortcut instead, though, which is Ctrl+Shift+I on a PC or Command+Option+I on a Mac.

Try it out! Insert a code chunk below this paragraph. Then, see if you can create an object called 'sum_10_20' that equals the sum of the integers from 10 to 20. Write 'sum_10_20' on the next line so that the result prints out when you run the code in the chunk or knit the document.



## Packages

Packages are like apps for R that expand its functionality. They need to be installed just once on your computer. As an example, you can install the `gapminder` package by typing `install.packages("gapminder")` into the console (bottom left quadrant of RStudio) and pressing enter on your keyboard. You should not include package installation code in an R Markdown file/code chunk, though, because then it would re-install the package each time you compiled the document.

Packages need to be loaded before each use. Usually, code that loads necessary packages is included in a code chunk at the top of an R Markdown file. We'll just load the `gapminder` package right here, though:

```{r}
# Loading the gapminder package (but not installing it here)

library(gapminder)
```



## Knitting

Afterwards, to knit (compile) the document, place your cursor anywhere in this editing window and press Ctrl+Shift+K on a PC or Command+Shift+K on a Mac. After a few seconds you should see the document pop up.



